Addressing selection bias and measurement error in COVID-19 case count data using auxiliary information

### 2022-03-30 write-up/letters/cover letter-covid.tex
paper is ready; tg

\documentclass[10pt]{letter} % Default font size of the document, change to 10pt to fit more text

@@

Kindly find enclosed the Report titled ``Statistical paradoxes in coronavirus case-counts: selection bias, measurement error, and the COVID-19 pandemic,'' by Walter Dempsey.

The report contains novel empirical findings and an investigation of statistical paradoxes arising from the analysis of reported coronavirus case-counts that are being used by governments to guide policies and decision making. We have discovered an estimator and statistical decomposition of the associated error that shows the 21.5 million tests conducted in the US provides information on COVID-19 prevalence equivalent to a simple random sample of size 92. Analogous decompositions for the rate of change in prevalence and the effective reproduction number are used to demonstrate how the magnitude and sign of estimation bias varies over time, a bias which has strong policy implications. Overall, our findings demonstrate the statistical limitations of increased testing in the presence of selection bias and measurement error as well as how easy it is for researchers analysing reported coronavirus case-counts to not only fool themselves but be confident in their incorrect conclusions. We believe the work presents an important advance in the statistical understanding of the COVID-19 pandemic.

>>>>

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text

@@

Kindly find enclosed the revised version of the manuscript number AOAS2006-048 titled ``Statistical paradoxes in coronavirus case-counts: selection bias, measurement error, and the COVID-19 pandemic'' now titled ``Addressing selection bias and measurement error in COVID-19 case count data using auxiliary information.''  We have revised the article to address the referees’ requests about exposition and other points of clarification.  The reports called for a major revision to make the manuscript ``forward compatible'', to re-structure the manuscript to present figures and tables near their corresponding discussion, to discuss potential non-response bias in the motivating studies, to include a more careful discussion of COVID-19 testing and disease dynamics, to include a more precise definition of the active infection rate and how it impacts model-based estimation, and to more carefully tie the proposal to related parts of the epidemiological literature. The new version complies with these requests.

We believe the work presents an important advance in how researchers should think about addressing time-varying selection bias and measurement error to better understanding the COVID-19 pandemic trajectory. Attached please find a detailed response to the referees outlining how the changes to the manuscript address the concerns raised in their reports.
